---
title: "Montreal"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

First, we load up the packages and set the API key.  

```{r}
library(cancensus)
library(tidyverse)
library(sf)
options(cancensus.api_key = "CensusMapper_6293253518cbfcfc86b62e83f056f92f")
```

"CA16" is the code for the 2016 Census; the CMA code for Montreal is 24462.  

We can browse possible variables with: 

```{r, eval = FALSE}
View(list_census_vectors("CA16"))
```

We want language spoken most often at home for several common languages in Quebec.  

This will get us the following codes for: 

* English: v_CA16_1364
* French: v_CA16_1367
* Arabic: v_CA16_1658
* Spanish: v_CA16_1958
* Italian: v_CA16_1949
* Haitian Creole: v_CA16_1736
* Greek: v_CA16_1883
* Chinese: v_CA16_2057
* Other: total (v_CA16_527) minus the rest
* Multiple: v_CA16_2147

And we grab the data by Census tract: 

```{r}
language <- c("v_CA16_1364", "v_CA16_1367", "v_CA16_1658", "v_CA16_1958", 
              "v_CA16_1949", "v_CA16_1736", "v_CA16_1883", "v_CA16_2057", 
              "v_CA16_527", "v_CA16_2147")

montreal <- get_census(dataset = "CA16", regions = list(CMA = "24462"), 
                       vectors = language, level = "CT", geo_format = "sf", 
                       labels = "short")
```

Let's write out to an RDS: 

```{r}
write_rds(montreal, "montreal.rds")
```

Next, we need to generate dots.  Let's try first for English speakers in Montreal.  

```{r}
english_dots <- montreal %>%
  st_transform(32618) %>%
  mutate(english = as.integer(v_CA16_1364 / 200)) %>%
  mutate(english = ifelse(is.na(english), 0, english)) %>%
  st_sample(., .$english) %>%
  st_sf()
```

And map them: 

```{r}
library(tmap)
library(tmaptools)

url <- "https://api.mapbox.com/styles/v1/kwalkertcu/cj0jov12u007n2sqppuxe3fvr/tiles/256/{z}/{x}/{y}?access_token=pk.eyJ1Ijoia3dhbGtlcnRjdSIsImEiOiJMRk9JSmRvIn0.l1y2jHZ6IARHM_rA1-X45A"

osm <- read_osm(bb(montreal), type = url)

b1 <- tm_shape(osm) + 
  tm_raster() + 
  tm_shape(english_dots) + 
  tm_dots(alpha = 0.4, col = "red") + 
  tm_layout(title = "English as first language in the Montreal area") + 
  tm_credits("1 dot = 200 people. Data source: 2016 Canadian Census via the R cancensus package")

save_tmap(b1, "img/english.jpg")

```

And French: 

```{r}

french_dots <- montreal %>%
  st_transform(32618) %>%
  mutate(french = as.integer(v_CA16_1367 / 200)) %>%
  mutate(french = ifelse(is.na(french), 0, french)) %>%
  st_sample(., .$french) %>%
  st_sf()

b2 <- tm_shape(osm) + 
  tm_raster() + 
  tm_shape(french_dots) + 
  tm_dots(alpha = 0.4, col = "#1C5BA2") + 
  tm_layout(title = "French as first language in the Montreal area") + 
  tm_credits("1 dot = 200 people. Data source: 2016 Canadian Census via the R cancensus package")

save_tmap(b2, "img/french.jpg")

```

From here, we'll be able to walk through, generate dots, and load to ArcGIS Online.  Let's write to a shapefile just in case though.  

* English: v_CA16_1364
* French: v_CA16_1367
* Arabic: v_CA16_1658
* Spanish: v_CA16_1958
* Italian: v_CA16_1949
* Haitian Creole: v_CA16_1736
* Greek: v_CA16_1883
* Chinese: v_CA16_2057
* Other: total (v_CA16_527) minus the rest
* Multiple: v_CA16_2147

```{r}
montreal_out <- montreal %>%
  select(GeoUID, english = v_CA16_1364, 
         french = v_CA16_1367, 
         arabic = v_CA16_1658, 
         spanish = v_CA16_1958, 
         italian = v_CA16_1949, 
         creole = v_CA16_1736, 
         greek = v_CA16_1883, 
         chinese = v_CA16_2057, 
         multiple = v_CA16_2147, 
         total = v_CA16_527) %>%
  mutate(other = total - (english + french + arabic + spanish + italian + creole + greek + chinese))
  

st_write(montreal_out, "data/montreal.shp", delete_layer = TRUE)
```

